---
title: "Final Project Manuscript"
subtitle: ""
author: Vinayak Revisanker
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github

---

Libraries

```{r}
library(tidyverse)
library(hms)
library(ggplot2)
```


Reading and loading CSV file
```{r}
methane <- read_csv("Methane in the Furnance.csv", show_col_types = FALSE)

methane

```


checking for missing values
```{r}
missing_values <- colSums(is.na(methane))

print(missing_values)

```


summary statistics
```{r}
co2_summary <- summary(methane$co2)
methane_summary <- summary(methane$methane)


print(co2_summary)
print(methane_summary)
```

Directories for plots and results
```{r}
dir.create("plots", showWarnings = FALSE)
dir.create("results", showWarnings = FALSE)
```


plot of CO2 levels over time
```{r}
co2_plot <- ggplot(methane, aes(x = obs, y = co2)) +
  geom_line(color = "blue") +
  labs(title = "CO2 Levels Over Time", x = "Time", y = "CO2 Level") +
  theme_minimal()
ggsave("plots/co2_plot.png", plot = co2_plot)

print(co2_plot)

```


Plot of methane levels over time
```{r}
methane_plot <- ggplot(methane, aes(x = obs, y = methane)) +
  geom_line(color = "orange") +
  labs(title = "Methane Levels Over Time", x = "Time", y = "Methane Level") +
  theme_minimal()
ggsave("plots/methane_plot.png", plot = methane_plot)

print(methane_plot)

```


Correlation between CO2 and methane levels

```{r}
correlation <- cor(methane$co2, methane$methane, use = "complete.obs")
print(paste("Correlation between CO2 and methane levels:", correlation))

```


Splitting data into training and testing sets
```{r}
set.seed(123)
train_indices <- sample(1:nrow(methane), size = 0.8 * nrow(methane))
train_data <- methane[train_indices, ]
test_data <- methane[-train_indices, ]

```



Linear model and summary
```{r}
lm_fit <- lm(co2 ~ methane, data = train_data)
lm_summary <- summary(lm_fit)
capture.output(lm_summary, file = "results/lm_summary.txt")

print(lm_fit)
print(lm_summary)
```
Model evaulation
```{r}
predictions <- predict(lm_fit, newdata = test_data)
actuals <- test_data$co2
residuals <- actuals - predictions
mse <- mean(residuals^2)
rmse <- sqrt(mse)
r_squared <- lm_summary$r.squared

print(paste("RMSE:", rmse))
print(paste("R-squared:", r_squared))


evaluation_metrics <- c(paste("RMSE:", rmse), paste("R-squared:", r_squared))
write(evaluation_metrics, file = "results/metrics_summary.txt")


```

Manuscript directory and results

```{r}
dir.create("manuscript", showWarnings = FALSE)



correlation <- readLines("results/analysis_summary.txt")
lm_summary <- readLines("results/lm_summary.txt")
metrics_summary <- readLines("results/metrics_summary.txt")


print(correlation)
print(lm_summary)
print(metrics_summary)
```

# Discussion

## Summary and Interpretation
_The analysis showed that both CO2 and methane levels fluctuate over time. The correlation between the two gases will provide insight into their relationship in the furnace environment._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_Understanding the behavior of methane and carbon dioxide in furnaces can help in developing better control strategies to reduce emissions. Further research with more context-specific data is recommended._



# References



>>>>>>> 0a8029a263c0a128bc84961770acd801f698db05
