---
title: "R Coding Exercise"
By: Vinayak Revisanker zuk198
editor: 
  markdown: 
    wrap: 72
---

# load dslabs package

library("dslabs") 

#get an overview of data structure str(gapminder)

# get a summary of data

summary(gapminder)

class(gapminder)

# Assign data from Africa to africadata

africadata <- subset(gapminder, continent == "Africa")

# overview of africadata

str(africadata)

# summary of africadata

summary(africadata)

# 2 new datasets

africadata_inf_life <- africadata[, c("infant_mortality", "life_expectancy")]
africadata_pop_life <- africadata[, c("population", "life_expectancy")]

# for africadata_inf_life

str(africadata_inf_life)
summary(africadata_inf_life)

# for africadata_inf_life

str(africadata_pop_life)
summary(africadata_pop_life)

# plotting data

#plotting for infant mortality and life expectancy
plot(africadata_inf_life$infant_mortality, africadata_inf_life$life_expectancy,
xlab = "Infant Mortality", ylab = "Life Expectancy", main = "Life
Expectancy vs. Infant Mortality")

plot(africadata_pop_life$population, africadata_pop_life$life_expectancy,
xlab = "Population (log scale)", ylab = "Life Expectancy", main = "Life
Expectancy vs. Population Size", log = "x")

# More data processing

missing_years <- table(africadata$year[is.na(africadata$infant_mortality)])
missing_years

# Extract data for the year 2000

africadata_2000 <- subset(africadata, year == 2000) 

#Overview and summary of the 2000 dataset
str(africadata_2000) 

#Display a summary of africadata_2000
 summary(africadata_2000) 
# More plotting

#plotting for infant mortality and life expectancy in the year 2000

plot(africadata_2000$infant_mortality, africadata_2000$life_expectancy,
xlab = "Infant Mortality", ylab = "Life Expectancy", main = "Life
Expectancy vs. Infant Mortality (2000)")

# Plot life expectancy and population size in the year 2000

plot(africadata_2000$population, africadata_2000$life_expectancy, xlab =
"Population (log scale)", ylab = "Life Expectancy", main = "Life
Expectancy vs. Population Size (2000)", log = "x")

# Simple model fits

# simple model fits for life expectancy and infant mortality based on the data from Africa from year 2000

model_fit1 <- lm(life_expectancy ~ infant_mortality, data = africadata_2000)
summary(model_fit1)

#simple model fits for life expectancy and population based on the data from Africa from year 2000 
model_fit2 <- lm(life_expectancy ~ population, data = africadata_2000)
summary(model_fit2)

# This section contributed by Brody Johnson

## More data exploration

In this section, we will explore the data set called "heights" which
contains two variables of the participants: self-reported gender and
height.

```{r}

str(heights) #structure of the data 
summary(heights) #summary statistics 
summary(heights$height[heights$sex=="Male"]) #summary statistics of height of males only
summary(heights$height[heights$sex=="Female"]) #summary statistics of height of females only 
class(heights) #data frame
```

Based on the structure and summary statistics, no entries appear to have
been entered incorrectly so no data cleaning is needed yet. Next, we
will plot these variables of interest to see their distributions. The
boxplots appear to be relatively normally distributed with several
outliers for each gender. The overlapping histograms are also
interesting in that it confirms they are both roughly noramlly
distributed and how many more male participants there were.

```{r}

plot(heights$sex, heights$height,
     xlab = "Gender", ylab = "Height", main = "Gender vs. Height")
```

```{r}
hist(heights$height[heights$sex=="Male"], col=rgb(0,0,1,0.25), xlim=c(min(heights$height), max(heights$height)), ylim=c(0,500), main="Male vs Female Height Histograms", xlab = "Height", ylab="Frequency")
hist(heights$height[heights$sex=="Female"], col=rgb(1,0,0,0.25),add=TRUE)
legend("topright", legend = c("Male", "Female"), fill = c(rgb(0,0,1,0.25), rgb(1,0,0,0.25)))
```

The linear model ran below confirms our suspicion that gender has an
effect on a persons height (p-value less than significance level of
0.05). That is, Males are roughly 4.37 inches taller than females.

```{r}
heightsfit = lm(height ~ sex, data = heights)
summary(heightsfit)
```
